---
title: "TakeHome_Ex02"
author: "TzuTing Hunag"
date: "February 25, 2025"
date-modified: "last-modified"
execute:
  warning: false
---

## Get Starting

```{r}
pacman::p_load(readxl, gifski, gapminder,
               plotly, gganimate,tidyverse, CGPfunctions, ggHoriPlot, tsibble, readxl, feasts, fable, seasonal, scales)
```

## Data Preparation

### Import Data

This code imports three datasets from an Excel file named **"Trade_Region_Market.xlsx"**, each from a different sheet:

-   **"T1"** → `import_m`: Contains import data

-   **"T2"** → `export_m`: Contains export data

-   **"T3"** → `re_export_m`: Contains re-export data

The `skip = 10` parameter ensures that the first **10 rows** are skipped, likely to remove headers or metadata, keeping only the relevant data.

```{r}
import_m <- read_excel("data/Trade_Region_Market.xlsx", sheet = "T1", skip = 10)
export_m <- read_excel("data/Trade_Region_Market.xlsx", sheet = "T2", skip = 10)
re_export_m <- read_excel("data/Trade_Region_Market.xlsx", sheet = "T3", skip = 10)

```

### Reshaping and Formatting Export/Import Data

-   **Pivoting Data**:

    -   Converts the dataset from a **wide format** to a **long format** using `pivot_longer()`.

    -   Keeps the **"Data Series"** column unchanged.

    -   Transforms all other columns (representing months and years) into two new columns:

        -   `"Month_Year"` → Contains the original column names (e.g., "2023 Jan").

        -   `"Export_Value"` → Stores the corresponding numerical export values.

-   **Date Conversion**:

    -   Uses `yearmonth(Month_Year)` to convert the `"Month_Year"` column into a proper **date format**, ensuring it is recognized as a **year-month** object.

```{r}
export_long <- export_m %>%
  pivot_longer(cols = -`Data Series`, names_to = "Month_Year", values_to = "Export_Value") %>%
  mutate(Date = yearmonth(Month_Year)) %>%
  select(`Data Series`, Date, Export_Value)
```

```{r}
import_long <- import_m %>%
  pivot_longer(cols = -`Data Series`, names_to = "Month_Year", values_to = "Import_Value") %>%
  mutate(Date = yearmonth(Month_Year)) %>%
  select(`Data Series`, Date, Import_Value) 

```

```{r}
colnames(import_m)
colnames(export_m)
colnames(re_export_m)
```

## TOTAL MERCHANDISE TRADE AT CURRENT PRICES, 2020 - 2024

### Export 

This code calculates **total exports** by summing **direct exports** and **re-exports** for each year from **2020 to 2024**:

1.  **Extract & Aggregate Data**:

    -   Converts `export_m` and `re_export_m` into long format.

    -   Filters **"Total All Markets"** and extracts the **year (`YYYY`)**.

    -   Aggregates **Annual Export** and **Annual Re-Export** separately.

2.  **Compute Total Exports**:

    -   Merges **exports and re-exports** using `left_join()`.

    -   Calculates **Total Export** as:

        **Total Export=Annual Export+Annual Re-Export**

```{r}
# 計算 Export 年度總和
export_summary <- export_m %>%
  pivot_longer(cols = -`Data Series`, names_to = "Month_Year", values_to = "Export_Value") %>%
  filter(`Data Series` == "Total All Markets") %>%  # 只選擇 "Total All Markets"
  mutate(Year = as.integer(substr(Month_Year, 1, 4))) %>%  # 擷取 "YYYY" 為年份
  filter(Year >= 2020 & Year <= 2024) %>% 
  group_by(Year) %>%
  summarise(Annual_Export = sum(Export_Value, na.rm = TRUE)) %>%  # 每年加總
  arrange(Year)

# 計算 Re-Export 年度總和
reexport_summary <- re_export_m %>%
  pivot_longer(cols = -`Data Series`, names_to = "Month_Year", values_to = "ReExport_Value") %>%
  filter(`Data Series` == "Total All Markets") %>%  # 只選擇 "Total All Markets"
  mutate(Year = as.integer(substr(Month_Year, 1, 4))) %>%  # 擷取 "YYYY" 為年份
  filter(Year >= 2020 & Year <= 2024) %>% 
  group_by(Year) %>%
  summarise(Annual_ReExport = sum(ReExport_Value, na.rm = TRUE)) %>%  # 每年加總
  arrange(Year)

# 合併 Export + Re-Export
export_total_summary <- export_summary %>%
  left_join(reexport_summary, by = "Year") %>%
  mutate(Total_Export = Annual_Export + Annual_ReExport)  # 計算總出口

print(export_total_summary)  # 確保結果正確


```

### Import

```{r}

import_summary <- import_m %>%
  pivot_longer(cols = -`Data Series`, names_to = "Month_Year", values_to = "Import_Value") %>%
  filter(`Data Series` == "Total All Markets") %>%  # 這裡要加 `%>%`
  mutate(Year = as.integer(substr(Month_Year, 1, 4))) %>%  # 確保 Year 正確提取
  filter(Year >= 2020 & Year <= 2024) %>%  
  group_by(Year) %>%
  summarise(Annual_Import = sum(Import_Value, na.rm = TRUE)) %>%
  arrange(Year)  

print(import_summary)

```

### Combine both Export and Import Data

```{r}

combined_summary <- export_total_summary %>%
  rename(Value = Total_Export) %>%  
  mutate(Type = "Export") %>%
  bind_rows(
    import_summary %>%
      rename(Value = Annual_Import) %>%
      mutate(Type = "Import")
  )

print(combined_summary)  # 確保數據正確


```

```{r}
# 計算 2024 年的 Total Merchandise Trade
trade_summary <- combined_summary %>%
  group_by(Year) %>%
  summarise(Total_Trade = sum(Value, na.rm = TRUE)) %>%
  arrange(Year)

print(trade_summary)  # 確保數據正確

```

```{r}
# 計算 2024 相較 2023 的增長率
increase_2024 <- (trade_summary$Total_Trade[trade_summary$Year == 2024] - 
                  trade_summary$Total_Trade[trade_summary$Year == 2023]) / 
                  trade_summary$Total_Trade[trade_summary$Year == 2023] * 100

print(paste0("Merchandise Trade Increased: ", round(increase_2024, 1), "% in 2024"))

```

```{r}
total_trade_summary <- total_trade_summary %>%
  mutate(Change = c(NA, diff(Total_Trade)),  # 計算變化值
         Change_Perc = c(NA, round(100 * diff(Total_Trade) / Total_Trade[-n()], 1)))  # 計算變化率（%）

```

```{r}
ggplot(combined_summary, aes(x = Year, y = Value, fill = interaction(Year, Type))) +
  geom_col(width = 0.6, position = position_dodge2(preserve = "single")) +  # Import & Export 並排
  geom_text(aes(label = paste0(round(Value / 1000, 1), "B")), 
            position = position_dodge2(width = 0.6), vjust = -0.5, size = 3, color = "black", fontface = "bold") +

  # 折線圖：顯示 Total Merchandise Trade
  geom_line(data = total_trade_summary, aes(x = Year, y = Total_Trade, group = 1), 
            color = "black", size = 1.2, linetype = "solid", inherit.aes = FALSE) +  
  geom_point(data = total_trade_summary, aes(x = Year, y = Total_Trade), 
             size = 4, color = "red", inherit.aes = FALSE) +  
  geom_text(data = total_trade_summary, aes(x = Year, y = Total_Trade, 
                                            label = paste0(round(Total_Trade / 1000, 1), "B")), 
            vjust = -1, size = 4, fontface = "bold", color = "black", inherit.aes = FALSE) +  

  # 在折線圖上標示變化值和變化率（不標第一年）
  geom_text(data = total_trade_summary %>% filter(!is.na(Change_Perc)), 
            aes(x = Year - 0.5, y = Total_Trade + 50, 
                label = paste0(ifelse(Change_Perc >= 0, "+", ""), Change_Perc, "%"), 
                color = as.factor(Change_Perc >= 0)),  # 確保顏色設定
            size = 4, fontface = "bold", inherit.aes = FALSE) +

  # **手動設定顏色：上升綠色，下降紅色**
  scale_color_manual(values = c("TRUE" = "blue", "FALSE" = "red")) +


  # 縮小 "Total Merchandise Trade" 標籤大小，並調整位置
  annotate("text", x = 2022, y = max(total_trade_summary$Total_Trade) * 1.06, 
           label = "Total Merchandise Trade", size = 2.5, fontface = "bold", hjust = -1.5, vjust = 7, color = "black") +

  theme_minimal(base_size = 12) +
  labs(title = "Annual Merchandise Trade (2020-2024)",
       x = "Year", y = "Value (Billion S$)", fill = "Trade Type") +
  scale_x_continuous(breaks = seq(2020, 2024, 1)) +  # 確保年份顯示正確
  scale_y_continuous(labels = comma_format(scale = 1/1000), expand = expansion(mult = c(0, 0.15))) +  
  scale_fill_manual(values = c("2020.Export" = "#E57373",   
                               "2020.Import" = "#FFCDD2",   
                               "2021.Export" = "#FBC02D",   
                               "2021.Import" = "#FFF176",   
                               "2022.Export" = "#64B5F6",   
                               "2022.Import" = "#BBDEFB",   
                               "2023.Export" = "#81C784",   
                               "2023.Import" = "#C8E6C9",   
                               "2024.Export" = "#BA68C8",   
                               "2024.Import" = "#E1BEE7")) +  
  theme(
    plot.title = element_text(face = "bold", hjust = 0.5, size = 20),
    axis.text.x = element_text(face = "bold", size = 12),
    legend.position = "top"
  )


```
