---
title: "In-Class_Ex07"
author: "TzuTing Hunag"
date: "February 25, 2025"
date-modified: "last-modified"
execute:
  warning: false
---

```{r}
pacman::p_load(tidyverse, tsibble, feasts, fable, seasonal)
```

```{r}
ts_data <- read_csv("In-class Exercise 7/visitor_arrivals_by_air.csv")
```

```{r}
ts_data$`Month-Year` <- dmy(ts_data$`Month-Year`)
```

```{r}
ts_data_ts <- ts(ts_data)
head(ts_data_ts)
```

```{r}
ts_tsibble <- ts_data %>%
  mutate(Month = yearmonth(`Month-Year`)) %>%
  as_tsibble(index = `Month`)
```

```{r}
ts_longer <- ts_data %>%
  pivot_longer(cols = c(2:34),
               names_to = "Country",
               values_to = "Arrivals")
```

```{r}
ts_longer %>%
  filter(Country == "Vietnam") %>%
  ggplot(aes(x = `Month-Year`, 
             y = Arrivals))+
  geom_line(size = 0.5)
```

```{r}
ggplot(data = ts_longer, 
       aes(x = `Month-Year`, 
           y = Arrivals,
           color = Country))+
  geom_line(size = 0.5) +
  theme(legend.position = "bottom", 
        legend.box.spacing = unit(0.5, "cm"))
```

```{r}
#| echo: false
#| fig-height: 12
ggplot(data = ts_longer, 
       aes(x = `Month-Year`, 
           y = Arrivals))+
  geom_line(size = 0.5) +
  facet_wrap(~ Country,
             ncol = 3,
             scales = "free_y") +
  theme_bw()
```

```{r}
tsibble_longer <- ts_tsibble %>%
  pivot_longer(cols = c(2:34),
               names_to = "Country",
               values_to = "Arrivals")
```

```{r}
tsibble_longer %>%
  filter(Country == "Italy" |
         Country == "Vietnam" |
         Country == "United Kingdom" |
         Country == "Germany") %>% 
  gg_season(Arrivals)
```

```{r}
tsibble_longer %>%
  filter(Country == "Vietnam" |
         Country == "Italy") %>% 
  gg_subseries(Arrivals)
```

```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam") %>%
  ACF(Arrivals) %>% 
  autoplot()
```

```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam") %>%
  PACF(Arrivals) %>% 
  autoplot()
```

V.S. Multiple

```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam" |
         `Country` == "Italy" |
         `Country` == "United Kingdom" |
         `Country` == "China") %>%
  ACF(Arrivals) %>%
  autoplot()
```

```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam" |
         `Country` == "Italy" |
         `Country` == "United Kingdom" |
         `Country` == "China") %>%
  PACF(Arrivals) %>%
  autoplot()
```

```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam") %>%
  gg_tsdisplay(Arrivals)
```

STL

```{r}
tsibble_longer %>%
  filter(`Country` == "Vietnam") %>%
  model(stl = STL(Arrivals)) %>%
  components() %>%
  autoplot()
```

### **19.7.1 Time Series Data Sampling**

```{r}
vietnam_ts <- tsibble_longer %>%
  filter(Country == "Vietnam") %>% 
  mutate(Type = if_else(
    `Month-Year` >= "2019-01-01", 
    "Hold-out", "Training"))
```

```{r}
vietnam_train <- vietnam_ts %>%
  filter(`Month-Year` < "2019-01-01")
```

### **19.7.2 Exploratory Data Analysis (EDA): Time Series Data**

```{r}
vietnam_train %>%
  model(stl = STL(Arrivals)) %>%
  components() %>%
  autoplot()
```

### **19.7.3 Fitting forecasting models**

#### 19.7.3.1 Fitting Exponential Smoothing State Space (ETS) Models: fable methods

```{r}
ETS(y ~ error(c("A", "M")) 
    + trend(c("N", "A", "Ad")) 
    + season(c("N", "A", "M")))
```

#### 19.7.3.2 Fitting a simple exponential smoothing (SES)

```{r}
fit_ses <- vietnam_train %>%
  model(ETS(Arrivals ~ error("A") 
            + trend("N") 
            + season("N")))
fit_ses
```

#### 19.7.3.7 Damped Trend methods

```{r}
vietnam_HAd <- vietnam_train %>%
  model(`Holt's method` = 
          ETS(Arrivals ~ error("A") +
                trend("Ad") + 
                season("N")))
vietnam_HAd %>% report()
```

```{r}
gg_tsresiduals(fit_ses)
```

### **19.7.4 Fitting ETS Methods with Season: Holt-Winters(WH)**

```{r}
Vietnam_WH <- vietnam_train %>%
  model(
    Additive = ETS(Arrivals ~ error("A") 
                   + trend("A") 
                   + season("A")),
    Multiplicative = ETS(Arrivals ~ error("M") 
                         + trend("A") 
                         + season("M"))
    )

Vietnam_WH %>% report()
```

### **19.7.5 Fitting multiple ETS Models**

```{r}
fit_ETS <- vietnam_train %>%
  model(`SES` = ETS(Arrivals ~ error("A") + 
                      trend("N") + 
                      season("N")),
        `Holt`= ETS(Arrivals ~ error("A") +
                      trend("A") +
                      season("N")),
        `damped Holt` = 
          ETS(Arrivals ~ error("A") +
                trend("Ad") + 
                season("N")),
        `WH_A` = ETS(
          Arrivals ~ error("A") + 
            trend("A") + 
            season("A")),
        `WH_M` = ETS(Arrivals ~ error("M") 
                         + trend("A") 
                         + season("M"))
  )
```

### **19.7.6 The model coefficient**

```{r}
fit_ETS %>%
  tidy()
```

### **19.7.7 Step 4: Model Comparison**

```{r}
fit_ETS %>% 
  report()
```

### **19.7.8 Step 5: Forecasting future values**

```{r}
fit_ETS %>%
  forecast(h = "12 months") %>%
  autoplot(vietnam_ts, 
           level = NULL)
```
